Versioned data: why it’s needed and how it can be achieved (easily and cheaply)

### 2017-08-03 ms.tex
reworked summary

Sharing data has become a key part of reproducible workflows in science.  To help this process, numerous linked tools have been developed to allow quick and easy data sharing.  Thus far, however, data has been largely considered a fixed and unchanging product---the particular data file for a particular analysis can be quickly unloaded, given a digital identifier, and shared.  While this is highly useful, this framework does not account for on-going scientific progress: for many important problems, epitomised by problems that lead to meta-analysis, datasets will continue to grow with time---more data points will be added, and new and more useful data structures will be created.  Essentially this means that data, like science in general, progress with time.  We suggest that a versioning system for data will facilitate this process.  Moreover, while the idea of versioned data has been taken up by large data sharing organizations, we present new tools for creating, archiving, and distributing versioned data easily, quickly and cheaply.  These new tools allow for individual research groups to shift from a static model of data curation to a dynamic and versioned model that naturally matches the scientific process.

>>>>

The sharing and re-use of data has become a cornerstone of modern science. To date, the curation and distribution of data has largely been conducted under one of two frameworks: data is either deposited in large, collaborative databases (often run by institutions and staffed by computational professionals) or else published as static files alongside a specific publication or as a standalone product released by a research group. For most medium-sized datasets, the former is clearly overkill and likely unwieldly and unsustainable and the latter, while convenient, is rather limited as it does not allow the dataset to improve nor to grow over time. This suggests is a substantial conceptual and technological gap in how we curate and share data. In this article, we argue that adopting the paradigm of versioned data, analogous to software versioning, can help fill this gap. We introduce a system called Versioned Data Delivery (VDD) and present novel computational tools to make creating, archiving, and distributing versioned data easily, quickly and cheaply. These new tools allow for individual research groups to shift from a static model of data curation to a dynamic and versioned model that naturally matches the scientific process.

%Sharing data has become a key part of reproducible workflows in science.  To help this process, numerous linked tools have been developed to allow quick and easy data sharing.  Thus far, however, data has been largely considered a fixed and unchanging product---the particular data file for a particular analysis can be quickly unloaded, given a digital identifier, and shared.  While this is highly useful, this framework does not account for on-going scientific progress: for many important problems, epitomised by problems that lead to meta-analysis, datasets will continue to grow with time---more data points will be added, and new and more useful data structures will be created.  Essentially this means that data, like science in general, progress with time.  We suggest that a versioning system for data will facilitate this process.  Moreover, while the idea of versioned data has been taken up by large data sharing organizations, we present new tools for creating, archiving, and distributing versioned data easily, quickly and cheaply.  These new tools allow for individual research groups to shift from a static model of data curation to a dynamic and versioned model that naturally matches the scientific process.
